Personalization and probabilities: impersonal propensities in online grocery shopping

### 2017-08-31 log.rmd
entered 89/89 edits

## Thu 20 Jul 2017 08:59:07 BST
- today is Section E; I have about 7 hour left to finish this; so will need to get that section done today.
- working hard on the probability stuff - wrote a section introducting Popper; reframed arules discussion.
- TODO - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;

@@

>>>>

## Thu 20 Jul 2017 08:59:07 BST
- today is Section E; I have about 7 hour left to finish this; so will need to get that section done today.
- working hard on the probability stuff - wrote a section introducing Popper; reframed arules discussion.
- TODO - short paragraph summarising the social theory of list ordering; make sure to add stuff on the ordering-remapping of order in the list;

@@

## Thu 31 Aug 2017 16:34:43 BST
- long gap here -- due to important other activities such as summer holidays;  
- found the 1989 Humby article on demographics -- argues for the need to add lifestage to that of neighbourhood. 
- TODO: add reference to [@Humby_1989]

### 2017-08-31 reconstruction.rmd
entered 89/89 edits

This paper philosophically reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model. The reconstruction is empirical in certain respects. It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders. Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.  It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists. It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time. It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.  Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.

@@

There are two ways of reading this statement. One is to see Popper as simply stating one of the consequencies of his physicalist understanding of probabilities as tendencies inherent in a situation. Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean. I would suggest that Patel's description of DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.

Several times, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned different attempts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model. A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice. The precise meaning of 'good', however, remains open to many different attempts, experiments and variations. I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things.

Patel, for instance, addressed the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket. A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation for an item that complements rather than duplicates items in their list. How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?  It could only do that if had some sense of the relation between items already in the list and the field of possible recommendations. The 'personal relevance' model, therefore, sought to include the relation between recommendations and items already ordered. The data scientists constructed a new feature from the data concerning how the substitutabilities and complementarities between product.  Taking all the baskets of items purchased, rather in the way that association rules does, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived. The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response'). A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.`

>>>>

This paper reconstructs [@Dewey_1957] an online grocery ordering system undergoing an upgrade from a demographically-based recommendation system to a 'personal relevance' model. The reconstruction is empirical in certain respects. It departs from an ethnographic moment: being a member of the audience at an industry/academic conference where a data scientist was describing how recommendations were created for online grocery orders. Starting from this highly mediated and fragmentary encounter with a recommender system and one of its designers, the paper explores the problem of personally relevant recommendations.  It draws on anthropological and sociological research concerning shopping and list-making to situate the recommender system amongst the social ordering practices of shopping and lists. It makes use of archaeological approaches, in the sense developed by Michel Foucault [@Foucault_1972] in order to identify and map important functional statements and practices configuring the knowledge and order generated by such systems over time. It conducts several small-scale code-based experiments in order to reconstruct, using widely available code resources resources such as API (Application Programmer Interfaces) and software libraries for machine learning, some prototypical elements of the system in question.  Throughout, it draws  on the philosophy of science and media theory to frame an engagement with data, algorithms and recommendations in terms of probabilities and their worldly enactments.

@@

There are two ways of reading this statement. One is to see Popper as stating a triusm,  the obvious consequence of his physicalist understanding of probabilities as tendencies inherent in a situation. Another is to focus on the pivotal phrase 'possibilities trying to realize themselves' by asking what 'trying to realize' might mean in practice. I would suggest that DunnHumby's work on the new personal relevance model is just such a 'trying to realize' in important ways.

Several times, Patel emphasized the importance of 'constructing good features' from the data, and much of her presentation concerned different attempts to construct 'good features.' A 'feature' in the context of machine learning refers to a variable included in a predictive model. A 'good feature' is one that somehow improves the accuracy, precision, specificity or any of the panoply of measures applied to machine learning models in practice.
HERE
The precise meaning of 'good', however, remains open to many different attempts, experiments and variations. I find it interesting that the main efforts that DunnHumby made to  construct good features were not closely associated with personhood, but with the relations -- the propensities -- between things. Patel, for instance, addressed the problem of 'basket similarity,' the problem that the recommender system might recommend something similar to items already in a customer's basket. A customer might be willing to substitute a similar item for something they have already chosen, but they are more likely to accept a recommendation that complements rather than duplicates items in their list. How could the recommender system sense the possible substitutions and complementaries between things for a given customer, especially since a person's sensibilities and susceptibilities concerning groceries will be subtly shaped by wider social groups, orderings and circumstances?  It could only do that if had some sense of the relation between items already in the list and the field of possible recommendations. The 'personal relevance' model, therefore, sought to include the relation between recommendations and items already ordered. The data scientists constructed a new feature from the data concerning how the substitutabilities and complementarities between product.  Taking all the baskets of items purchased, rather in the way that association rules does, a new feature, what DunnyHumby termed 'self-learning substitutes' was derived. The 'self learning substitutes' feature of the relevancy model draws on another data matrix, the product similarity matrix, itself generated from all previous recommendations that have led to conversion events (that is, 'a user clicked as the response'). A customer's recommendation list might already include `Tesco Ginger Nuts 300g.`   The 'self-learning substitution model' seeks to replace list items with similar but higher value items: `Mcvitie's Ginger Moments 400g.`
